---
title: "Jenny Bryan Webinar on `reprex`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
library(reprex)
```

>
"Making a good reprex is difficult,
you're asking others (me) to experience your pain.
A good reprex is how you meet them halfway." - Jenny Bryan
>

# Intro: resources

* [RStudio Webinars](https://resources.rstudio.com/webinars){target="_blank"}
* [dos and donts of `reprex`](https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html){target="_blank"}
* [`reprex` on Github](https://github.com/tidyverse/reprex){target="_blank"}
* [Jenny Bryan's Slide Deck](https://speakerdeck.com/jennybc/reprex-reproducible-examples-with-r){target="_blank"}
* Current `reprex` v`r packageVersion('reprex')`
  

# How to generate a `reprex`

1. Clipboard -> `reprex::reprex()`
2. RStudio addin:
   i) launch GUI
   i) selection -> key binding (Crtl+Shift+R)
3. An inline expression
4. A separate file
5. **note**: `Rmarkdown` possible for paste into e-mail:
   * reprex -> highlight -> copy -> paste

## Examples
```{r examples, eval = FALSE}
# clipboard
x <- factor("a")
y <- factor("b")
c(x, y)      # what happened here?

# inline
reprex({
  x <- factor("a")
  y <- factor("b")
  c(x, y)
})

# external file
reprex(input = "my_reprex.R")
```

# How to make a good `reprex`
The `reprex` philosophy:

|     Tennet     |          example           |
|:--------------:|:-------------------------- |
| code that actually runs             | **it must be self contained** |
| code the reader doesn't have to run | **keep it simple** |
| code the reader could easily run    | **NEVER SCREEN SHOT!** |



## Code that runs
```{r praise, eval = FALSE}
# attempt 1
template <- "${ESCLAMATION} - your code is ${adjective}"
praise(template)

# attempt 2
library(praise)
praise(template)

# success!
library(praise)
template <- "${ESCLAMATION} - your code is ${adjective}"
praise(template)
```


## Simplist possible example
```{r}
# KISS -> a 'minimal' example
read.csv(text = "a,b\n1,2\n3,4")  # inline creation of df
```


## Code Easy to run
```{r}
# Do not copy/paste from the R console!
# Difficult fiddly edits to reproduce.
test <- "'Header\nLine 2'\nValue"
cat(test)
readr::read_csv(test)   # `reprex::reprex_clean()`
```

```{r clean, eval = FALSE}
# removes `#` at lines
reprex::reprex::clean()
```




## You can include plots as well!
Figure uploaded to `imgur.com`:
```{r plots, eval = FALSE}
gapminder %>% 
  dplyr::filter(continent != "Oceania") %>% 
  ggplot(aes(x = year, y = lifeExp, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) +
  facet_wrap(~ continent) +
  scale_color_manual(values = country_colors) +
  theme_bw() + theme(strip.text = element_text(size = rel(1.1))) +
  NULL
```



# Output options
```{r output, eval = FALSE}
# standard
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
})


# You can send output files to directory
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, outfile = "my_corr")


# Include session info
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, si = TRUE)


# Re-style your code
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, style = TRUE)    # tidyverse style


# change output format for venue
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, venue = "so")   # options: gh <def>, so, ds, r, rtf


# capture standard error out
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, std_out_err = TRUE)
```




# Jen’s Yellowstone Bison Say
```{r, echo = FALSE}
cowsay::say(what = "Hello Jen! Thanks for preserving my genetics!",
            by = "buffalo")
```

