---
title: "Jenny Bryan Webinar on `reprex`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r echo = FALSE}
library(reprex)
```

# Resources
* [https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html](reprex dos and donts)
* [https://github.com/tidyverse/reprex](reprex github)
* [rstd.io/reprex](info materials)
* Current `reprex` v`r packageVersion('reprex')`
  

# How to generate a `reprex`
1. from clipboard -> `reprex::reprex()`
2. use addin -> RRStudio
   i) launch GUI
   i) selection -> key binding (Crtl+Shift+R)
3. use an inline expression
4. use a separate file
5. NOTE: reprex -> highlight -> copy -> paste -> rmarkdown


# How to make a good `reprex`
The `reprex` philosophy:

* code that runs                       -> it must be self contained
* code that reader doesn't have to run -> keep it simple
* code that reader could easily run    -> NEVER SCREEN SHOT!!!!



## Example:
```{r}
x <- factor("a")
y <- factor("b")
c(x, y)
```



## Simplist possible example
```{r}
read.csv(text = "a,b\n1,2\n3,4")  # inline creation of df
```






## You can include plots as well!
```{r, eval = FALSE}
reprex({
  library(magrittr)
  library(ggplot2)
  set.seed(101)
  data.frame(x = 1:100,
             y = 1:100 + rnorm(100, sd = 10)) %>% 
    ggplot2::ggplot(ggplot2::aes(x = x, y = y)) +
    ggplot2::geom_point() +
    ggplot2::geom_smooth(method = "loess") +
    NULL
})
```


```{r, eval = FALSE}
# copy all code
# removes `#` at lines
reprex::reprex::clean()
```


# Output options
```{r, eval = FALSE}
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
})
```


## You can send output files to directory
```{r, eval = FALSE}
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, outfile = "my_corr")
```


## Include session info
```{r, eval = FALSE}
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, si = TRUE)    # session info
```


## Re-style your code
```{r, eval = FALSE}
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, style = TRUE)    # tidyverse style
```



## capture standard error out
```{r eval = FALSE}
reprex({
  x <- rnorm(100)
  y <- rnorm(100)
  stats::cor(x, y)
}, std_out_err = TRUE)
```




# Jen’s Yellowstone Bison Say
```{r, echo = FALSE}
cowsay::say(what = "Hello Jen! Thanks for preserving my genetics!",
            by = "buffalo")
```

